\bf Model Visualization Techniques for a Social Network Model

### 2017-05-10 paper.Rnw
add siena g-o-f description and plots

The \texttt{RSiena} software that fits the models to data also comes equipped with the \verbatim{sienaGOF()} function for examining model fit. This function ``assess[es] the fit of the model with respect to auxiliary statistics of networks" \citep[p.~53]{RSienamanual}. Examples of the auxiliary statistics are the out- or indegree distribution on the nodes, with the option for users to create their own.

@@

# running Siena
#friendship <- sienaDependent(friendshipData)
#alcohol <- varCovar(drink[20:35,])
#mydata <- sienaDataCreate(friendship, alcohol)
#myeffnull <- getEffects(mydata)
#myalgorithm <- sienaAlgorithmCreate(projname = 's16_3')
set.seed(823746)
load("data/ansnullpaper.rda")
ansnullchains <- get_chain_info(ansnull)
ansnullchains %>%
  filter(period == 1) %>%  #only look at chains from wave 1 to wave 2
  group_by(rep) %>%
  select(rep, from = ego, to = alter) %>%
  mutate(val = as.numeric(!from == to),
         from = paste0("V", parse_number(from)+1), # make the chains
         to = paste0("V", parse_number(to)+1)) -> ansnullchainsw1w2
colnames(fd2.w1) <- paste0("V", 1:16)
rownames(fd2.w1) <- paste0("V", 1:16)
wave1friends <- fortify(as.adjmat(fd2.w1))
ms1 <- listMicrosteps(dat = wave1friends,
                      microsteps = filter(ansnullchainsw1w2, rep == 1))
microsteps <- ms1[1:20]

>>>>

The \texttt{RSiena} software that fits the models to data also comes equipped with the \verbatim{sienaGOF()} function for examining model fit. This function ``assess[es] the fit of the model with respect to auxiliary statistics of networks" \citep[p.~53]{RSienamanual}. Examples of the auxiliary statistics are the out- or indegree distribution on the nodes, with the option for users to input their own statistics to examine. The fit is tested using Mahalanobis distance to determine a $p$-value when comparing the statistics calculated from the observed data (excluding the first observation that is conditioned on) to the distribution of the statistics when calculated on the simulated waves. So, good fit is acheived when $p$ is large. XX more detail about mahalanobis distance and calculating p-value XX

The \verbatim{plot.sienaGOF()} function allows us to visualize this fit. Below is an example of what the goodness-of-fit plot output looks like for indegree and outdegree statistics for a small model. Box plots and violin plots are drawn on top of each other in Figure~\ref{fig:sienagof} in order to show the distribution of the simulated network observations compared to the true data shown in red points connected by red lines. If the red points lie well within the simulated values, the model is a good fit to the data. The model examined in Figure~\ref{fig:sienagof} appears to do a better job of capturing the outdegree distribution than the indegree distrubtion of the data.

<<sienagof, fig.cap="Goodness of fit measures for two network statistics: indegree and outdegree cumulative distributions. On the x-axis is the degree value, and on the y-axis is the number of times that degree value occurs in the data.">>=
load("data/ansnullpaper.rda")
gof1 <- sienaGOF(ansnull, OutdegreeDistribution, varName ="friendship")
gof2 <- sienaGOF(ansnull, IndegreeDistribution, varName ="friendship")
p1 <- plot(gof1)
p2 <- plot(gof2)
grid.arrange(p1, p2)
@

@@

# running Siena
#friendship <- sienaDependent(friendshipData)
#alcohol <- varCovar(drink[20:35,])
#mydata <- sienaDataCreate(friendship, alcohol)
#myeffnull <- getEffects(mydata)
#myalgorithm <- sienaAlgorithmCreate(projname = 's16_3')
set.seed(823746)
ansnullchains <- get_chain_info(ansnull)
ansnullchains %>%
  filter(period == 1) %>%  #only look at chains from wave 1 to wave 2
  group_by(rep) %>%
  select(rep, from = ego, to = alter) %>%
  mutate(val = as.numeric(!from == to),
         from = paste0("V", parse_number(from)+1), # make the chains
         to = paste0("V", parse_number(to)+1)) -> ansnullchainsw1w2
colnames(fd2.w1) <- paste0("V", 1:16)
rownames(fd2.w1) <- paste0("V", 1:16)
wave1friends <- fortify(as.adjmat(fd2.w1))
ms1 <- listMicrosteps(dat = wave1friends,
                      microsteps = filter(ansnullchainsw1w2, rep == 1))
microsteps <- ms1[1:20]
